---
title: "check_annotations"
author: "Nico Bast"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    theme: cerulean
    code_folding: hide
  word_document:
    toc: yes
editor_options:
  chunk_output_type: console
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

This R Markdown Shiny document takes an annotation export from file and generated an output that provides a technical check of the annotations  

## Select a file

```{r file selection}

#UI to select file
ui <- fluidPage(
  titlePanel("Select a File"),
  sidebarLayout(
    sidebarPanel(
      fileInput("file", "Choose a file")
    ),
    mainPanel(
      textOutput("fileName"),
      tableOutput("fileContents")
    )
  )
)

#use function to read file
server <- function(input, output, session) {
  values <- reactiveValues(data = NULL)
  
  observeEvent(input$file, {
    req(input$file)
    values$data <- read.csv(input$file$datapath,sep=';',dec=',')
  })
  
  output$fileName <- renderText({
    if (is.null(input$file)) {
      return("No file selected")
    }
    paste("Selected file:", input$file$name)
  })
  
  output$fileContents <- renderTable({
    values$data
  })
}

shinyApp(ui = ui, server = server)

```

## Convert data
```{r}

# Access the file data stored in the reactive object
rater1 <- isolate(values$data)

# Perform operations on the file data
if (!is.null(fileData)) {
    #convert time data - consider whether time is provided as POXIXCT character or numeric character
  if(grepl(':',rater1$Onset_Time[1])){
    rater1$Offset_Time<-fun_timeconv2(rater1$Offset_Time)
    rater1$Onset_Time<-fun_timeconv2(rater1$Onset_Time)
  } else {
    rater1$Offset_Time<-as.numeric(rater1$Offset_Time)
    rater1$Onset_Time<-as.numeric(rater1$Onset_Time)
  }
  head(rater1$Onset_Time)
} else {
  "No data available"
}


```



## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```



