---
title: "IRR_video_annotations"
author: "Nico Bast"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    theme: cerulean
    code_folding: hide
  word_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #show code in markdown

## packages
require(openxlsx)
require(ggplot2)
require(psych) #IRR metric


theme_set(theme_bw())

## paths
path<-'C:/Users/nico/PowerFolders/project_dynamic/video_annotations'

##required functions
###time conversion
fun_timeconv<-function(x){
  as.POSIXct(x,format = "%H:%M:%OS") #consider fractional seconds = milliseconds with %OS
}

###task selection function
fun_seltask<-function(x,var_name){
x<-x[,names(x) %in% c('Number','Onset_Time','Offset_Time',var_name)]
}
```

# read and select data

- reads INTERACT output for two rater and selects a specific sequences of a task
- manual edits:
- selct the rater in code (Rosa, Merlinda, Paula)
- select the rating category (Mutual, Avert, Refer, Follow, Share, Supporting.gesture, Supporting.behavior, Transcript)
- select the time frame of the benchmark sequence (< 5 min, > 5 min)


```{r read and select data}

rating_category<-'Share'
cutoff<-as.POSIXct("00:05:00",format = "%H:%M:%OS") #time frame

#select rater data
rater2<-read.xlsx(paste0(path,'/data/FRA_038_T5_ESCS_Übereinstimmung.xlsx'),sheet = 2) #MERLINDA
rater1<-read.xlsx(paste0(path,'/data/FRA_038_T5_ESCS_Übereinstimmung.xlsx'),sheet = 1,rows = 1:309) #ROSA

#convert time data
rater1$Offset_Time<-fun_timeconv(rater1$Offset_Time)
rater1$Onset_Time<-fun_timeconv(rater1$Onset_Time)
rater2$Offset_Time<-fun_timeconv(rater2$Offset_Time)
rater2$Onset_Time<-fun_timeconv(rater2$Onset_Time)

#select a task based on timestamp
rater1_task1<-rater1[rater1$Onset_Time<cutoff,]
rater2_task1<-rater2[rater2$Onset_Time<cutoff,]

#select a rating category
rater1_task1_category<-fun_seltask(rater1_task1,rating_category)
rater2_task1_category<-fun_seltask(rater2_task1,rating_category)

```

# Plot annotations for a category

Figure shows the number of annotated gaze events (x-axis) and onset and offset of that event (y-axis).
Overlap in Ratings can be done by looking at horizontal overlap at any given timepoint.

displayed rating category: `r rating_category`


```{r pressure, echo=FALSE, fig.height=10}

table(rater1_task1_category[,4])
table(rater2_task1_category[,4])

#combine onset and offset data with rater for plotting
y1<-rater1_task1_category[!is.na(rater1_task1_category[,4]),2:3]
y1$rater<-"Rosa"
y1$seq<-1:nrow(y1)
y2<-rater2_task1_category[!is.na(rater2_task1_category[,4]),2:3]
y2$rater<-"Merlinda"
y2$seq<-1:nrow(y2)
df<-rbind(y1,y2)


ggplot(df,aes(x=Onset_Time,y=Offset_Time,group=rater,color=rater))+geom_point(alpha=0.5)

ggplot(df) +
  geom_segment(aes(x = seq, xend = seq, y = Onset_Time, yend = Offset_Time,color=rater,linewidth=3)) +
  labs(title = "Bars with Onset and Offset on Y-axis",
       x = paste("gaze event number (gaze event:",rating_category,")"),
       y = "segment duration (min)")
  
```

# Create Matrix for calculating Inter-Rater-Reliability

```{r IRR matrix}

# convert to matrix to define inter-rater reliability

##create a matrix - function takes interact ratings and creates vector in tenth of second format with according rating label
fun_rater_vector<-function(rating_df){

###get max time
max_time<-max(rating_df$Offset_Time)
max_time<-format(max_time,"%H:%M:%OS2") #backconverted to character

###calculate the number of tens of seconds 
vector_rows<-as.numeric(substr(max_time,1,2))*360000+
  as.numeric(substr(max_time,4,5))*6000+
  as.numeric(substr(max_time,7,8))*100+
  as.numeric(substr(max_time,10,11))

#length of rater vector based on rated time
rater_vector<-rep(NA,vector_rows)

#function that fills a vector with categories based on ratings
for(i in 1:nrow(rating_df)){
start_row<-round(as.numeric(difftime(rating_df$Onset_Time[i],min(rating_df$Onset_Time),units='secs'))*100)
end_row<-round(as.numeric(difftime(rating_df$Offset_Time[i],min(rating_df$Onset_Time),units='secs'))*100)
category<-rating_df[i,4] 
if(is.na(category)){category<-'NONE'}

#print(c(i,start_row,end_row,category)) #DEBUGGING

rater_vector[start_row:end_row]<-category
}

return(rater_vector)
  
}

#convert to vector
test_vector<-fun_rater_vector(rating_df=rater1_task1_category)
test2_vector<-fun_rater_vector(rating_df=rater2_task1_category)

table(test_vector)
table(test2_vector)

#create a matrix combing ratings, consider different lengths
rating_df<-data.frame(matrix(NA, nrow=max(c(length(test_vector),length(test2_vector))),ncol=2))
rating_df$X1[1:length(test_vector)]<-test_vector
rating_df$X2[1:length(test2_vector)]<-test2_vector

table(rating_df[,1],rating_df[,2]) #look into contingency table

cohen.kappa(rating_df)

```

